---
title: "Superstore Retail Analytics"
author: Tay Chee Hsian
format: html
---

Load required packages.

```{r}
library(readxl)
library(dplyr)
```

Read the dataset.

```{r}
dataset <- read_excel("superstore.xlsx")
summary(dataset)
```

## Problem 1

```{r}
records <- dataset %>% 
  # Country that appear more than 200 times
  group_by(Country) %>% 
  filter(n() > 200) %>% 
  # Above that country's own shipping cost
  filter(Shipping.Cost > mean(Shipping.Cost, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # Profit for every unit of shipping cost
  mutate(Profit_Ratio = Profit / Shipping.Cost) %>% 
  # Sort Profit_Ratio in descending order
  arrange(desc(Profit_Ratio)) %>% 
  # Select specific attributes
  select(Customer.Name, Product.Name, Country, Sales, Profit, Shipping.Cost, Profit_Ratio) %>% 
  # Show top 15 records
  head(15)

records
```

## Problem 2

```{r}
order_analysis <- dataset %>% 
  # Extract each order details from every market-year combination
  group_by(Market, Year, Order.ID) %>% 
  # Compute sales and profit for each order
  summarise(
    Order_Sales = sum(Sales, na.rm = TRUE),
    Order_Profit = sum(Profit, na.rm = TRUE),
    .groups = "drop") %>% 
  # Discard those negative order sales.
  filter(Order_Sales > 0)

# Compute overall median profit
profit_med = median(order_analysis$Order_Profit, na.rm = TRUE)

market_analysis <- order_analysis %>% 
  # Merket-Year combination
  group_by(Market, Year) %>% 
  # Compute average profit and total sales for each order in market-year context
  summarise(
    Avg_Profit = mean(Order_Profit, na.rm = TRUE),
    Ttl_Sales = sum(Order_Sales, na.rm = TRUE),
    Orders = n(),
    .groups = "drop") %>% 
  # Discard number of orders is less than 50
  filter(Orders >= 50) %>% 
  # Present the result in market context
  group_by(Market) %>% 
  # Keep markets where no year drops below the median profit
  filter(all(Avg_Profit >= profit_med)) %>% 
  # Compute average sales across the remaining valid orders
  summarise(
    Avg_Sales = sum(Ttl_Sales) / sum(Orders),
    .groups = "drop") %>% 
  # Sort in descending order
  arrange(desc(Avg_Sales)) %>% 
  # Show first 10 rows
  head(10)

market_analysis
```

## Problem 3

```{r}
combinations <- dataset %>% 
  # Classify different sub category
  group_by(Sub.Category) %>% 
  # Define typical discount level using median
  mutate(Typical_Disc_Lvl = median(Discount, na.rm = TRUE)) %>% 
  # Filter suspicious discount that is greater than typical discount level multiply by 1.5
  # Multiply 1.5 for capturing unusually high discount
  filter(Discount > (Typical_Disc_Lvl * 1.5)) %>% 
  # Divide which suspicious discount actually loss making and non loss making
  mutate(Profit_Status = ifelse(Profit > 0, "non-loss-making", "loss-making")) %>% 
  # Focus on Market Segment combination
  group_by(Market, Segment) %>% 
  # For each combination, compute the number of non loss making and loss making
  summarise(
    Non_Loss_Making_Count = sum(Profit_Status == "non-loss-making"),
    Loss_Making_Count = sum(Profit_Status == "loss-making"),
    Ttl_Suspicious = n(),
    .groups = "drop"
  ) %>% 
  # Compute the percentage of loss making in total suspicous record
  mutate(Loss_Making_Share = Loss_Making_Count / Ttl_Suspicious) %>% 
  # Preserve loss making share that over 40%
  filter(Loss_Making_Share > 0.4) %>% 
  # Sort in descending order
  arrange(desc(Loss_Making_Share))

combinations
```

## Problem 4

```{r}
# First condition: Quantity >= 2
filter_qty <- dataset %>% filter(Quantity >= 2)

ttl_profit_1 <- 0
row_count_1 <- 0
for(i in 1:nrow(filter_qty)) {
  ttl_profit_1 <- ttl_profit_1 + filter_qty$Profit[i]
  if(ttl_profit_1 > 25000) {
    row_count_1 <- i
    break
  }
}

record_1 <- filter_qty %>% 
  slice(row_count_1) %>% 
  mutate(Row = row_count_1, Cumulative_Profit = ttl_profit_1) %>% 
  select(Row, Order.ID, Customer.Name, Cumulative_Profit)

# Second condition: Quantity >= 2 and sort the sales from highest to lowest
reorder <- filter_qty %>% arrange(desc(Sales)) 

ttl_profit_2 <- 0
row_count_2 <- 0
for(i in 1:nrow(reorder)) {
  ttl_profit_2 <- ttl_profit_2 + reorder$Profit[i]
  if(ttl_profit_2 > 25000) {
    row_count_2 <- i
    break
  }
}

record_2 <- reorder %>% 
  slice(row_count_2) %>% 
  mutate(Row = row_count_2, Cumulative_Profit = ttl_profit_2) %>% 
  select(Row, Order.ID, Customer.Name, Cumulative_Profit)

record_1
record_2
```

The comparison shown that, after sorting by sales in descending order, the order processing volume required to reach the target profit of 25,000 was reduced from 776 rows to 6 rows. This demonstrates that the strategy of prior focusing on high profit orders can significantly affect the entire financial target achievement cycle. Even after filtering those orders that quantity less than 2, the profit contribution from large orders is still sufficient to allow the company to achieve its target while processing fewer orders.

## Problem 5

```{r}
resilient_prod <- dataset %>% 
  # Filter those products sold in at least 4 different countries
  group_by(Product.Name) %>% 
  filter(n_distinct(Country) >= 4) %>% 
  # Compute median discount for each product
  mutate(Med_Disc = median(Discount, na.rm = TRUE)) %>% 
  # Compute average profit that discount above the product's own median and below the median
  summarise(High_Avg_Profit = mean(Profit[Discount > Med_Disc]),
            Low_Avg_Profit = mean(Profit[Discount <= Med_Disc]),
            .groups = "drop") %>% 
  # Compute the gap between previous two conditions
  mutate(Gap = High_Avg_Profit - Low_Avg_Profit) %>% 
  # Preserve those products with high discount but still performs better
  filter(Gap > 0) %>% 
  # Sort in descending order
  arrange(desc(Gap)) %>% 
  # Show the first 5 rows
  head(5)

resilient_prod
```

## Problem 6

```{r}
market_city_shares <- dataset %>% 
  group_by(Market, City, Segment) %>% 
  # Compute city sales for each segment within each market
  summarise(
    Ttl_City_Sales = sum(Sales, na.rm = TRUE),
    .groups = "drop") %>% 
  group_by(Market, Segment) %>% 
  # Compute city share of its market sales
  mutate(
    Ttl_Market_Sales = sum(Ttl_City_Sales, na.rm = TRUE),
    Market_City_Shares = Ttl_City_Sales / Ttl_Market_Sales) %>% 
  # Find the city greatest share within each market
  filter(Market_City_Shares == max(Market_City_Shares, na.rm = TRUE)) %>% 
  # Keep cities contribute more than one third of market sales
  filter(Market_City_Shares > 1/3) %>% 
  # Show dominant city, market, segment, and shares
  select(City, Market, Segment, Market_City_Shares) %>% 
  # Sort shares in descending order
  arrange(desc(Market_City_Shares))

market_city_shares
```

## Problem 7

```{r}
finance_review <- dataset %>% 
  # Only final quarter (Oct to Dec) for each year
  filter(quarters(Order.Date) == "Q4") %>% 
  # Compute custom score
  mutate(Score = Profit - Shipping.Cost - (Sales * Discount)) %>% 
  # Group by Region, Year and Product to find product that appears most
  group_by(Region, Year, Product.Name) %>% 
  # Compute total score1
  summarise(Ttl_Score = sum(Score, na.rm = TRUE),
            .groups = "drop_last") %>% 
  # Only top-scoring product for each year within region
  filter(Ttl_Score == max(Ttl_Score)) %>% 
  ungroup() %>% 
  # Count which product appearing most
  count(Region, Product.Name, name = "Popular_Product") %>% 
  # Only top-scoring within the region
  group_by(Region) %>% 
  filter(Popular_Product == max(Popular_Product))

View(finance_review)
```

## Problem 8

```{r}
order_priority <- dataset %>% 
  # Compute the differences between order date and ship date in days
  mutate(ship_days = as.numeric(difftime(Ship.Date, Order.Date, units = "days"))) %>% 
  # Discard those NA ship day and less than 0
  filter(!is.na(ship_days), ship_days >= 0) %>% 
  # Compute average ship days across ship mode and order priority
  group_by(Ship.Mode, Order.Priority) %>% 
  summarise(
    Avg_Days = mean(ship_days, na.rm = TRUE),
    .groups = "drop"
  )

abnormal_cases <- order_priority %>% 
  group_by(Ship.Mode) %>% 
  # Compute cases where lower urgency unexpectedly ships faster than higher urgency within the same ship mode
  summarise(surprising_cases = 
      # low < medium
      sum(Avg_Days[Order.Priority == "Low"] < Avg_Days[Order.Priority == "Medium"], na.rm = TRUE) +
      # low < high
      sum(Avg_Days[Order.Priority == "Low"] < Avg_Days[Order.Priority == "High"], na.rm = TRUE) +
      # low < critical
      sum(Avg_Days[Order.Priority == "Low"] < Avg_Days[Order.Priority == "Critical"], na.rm = TRUE) +
      # medium < high
      sum(Avg_Days[Order.Priority == "Medium"] < Avg_Days[Order.Priority == "High"], na.rm = TRUE) +
      # medium < critical
      sum(Avg_Days[Order.Priority == "Medium"] < Avg_Days[Order.Priority == "Critical"], na.rm = TRUE) + 
      # high < critical
      sum(Avg_Days[Order.Priority == "High"] < Avg_Days[Order.Priority == "Critical"], na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  # Sort in descending order
  arrange(desc(surprising_cases))

abnormal_cases
```

## Problem 9

```{r}
cumulative_sales <- dataset %>% 
  # Scanning week by week within each year
  group_by(Year, weeknum) %>% 
  # Compute weekly sales and weekly profit
  summarise(
    Weekly_Sales = sum(Sales, na.rm = TRUE),
    Weekly_Profit = sum(Profit, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  # Ignore negative weekly sales
  filter(Weekly_Profit >= 0) %>% 
  # Sort year and weeknum in ascending order before calculating cumulative sales
  arrange(Year, weeknum) %>% 
  # Show yearly report
  group_by(Year) %>% 
  # Compute threshold and cumulative sales
  mutate(
    Threshold_Sales = sum(Weekly_Sales, na.rm = TRUE) * 0.2,
    Cumulative_Sales = cumsum(Weekly_Sales),
    Week_Count = row_number()
  ) %>% 
  # Filter cumulative sales that greater than threshold
  filter(Cumulative_Sales > Threshold_Sales) %>% 
  # Choose the first exceed threshold record
  slice_head(n = 1) %>% 
  ungroup() %>% 
  # Select year, triggering week, and cumulative sales as report
  select(Year, Triggering_Week = weeknum, Cumulative_Sales, Week_Count)

cumulative_sales
```

## Problem 10

```{r}
metric_calculation <- dataset %>% 
  # State and Sub.Category combinations
  # Region is grouped because it is ranked at the end
  group_by(Region, State, Sub.Category) %>% 
  # Compute the number of orders across state, sub.category, and region
  # Preprocess total profit, sales, shipping cost, and quantity for calculate new metrics
  summarise(Order_Count = n(),
            Ttl_Profit = sum(Profit, na.rm = TRUE),
            Ttl_Sales = sum(Sales, na.rm = TRUE),
            Ttl_Ship_Cost = sum(Shipping.Cost, na.rm = TRUE),
            Ttl_Qty = sum(Quantity, na.rm = TRUE),
            .groups = "drop") %>% 
  # Keep only combinations with at least 30 rows
  filter(Order_Count >= 30) %>% 
  # Compute new metrics: Profitability and Shipping Cost Per Unit
  mutate(Profitability = Ttl_Profit / Ttl_Sales,
         Ship_Cost_Per_Unit = Ttl_Ship_Cost / Ttl_Qty)

# Dataset wide average
Avg_Profitability <- mean(metric_calculation$Profitability)
Avg_Ship_Cost <- mean(metric_calculation$Ship_Cost_Per_Unit)

quiet_underperformers <- metric_calculation %>% 
  # Filter profitability belows the dataset wide average
  # Filter shipping cost per unit aboves the dataset wide average
  filter(Profitability < Avg_Profitability,
         Ship_Cost_Per_Unit > Avg_Ship_Cost) %>% 
  # Present result in region
  group_by(Region) %>% 
  # Sort in ascending order to find the worst profitability
  arrange(Profitability) %>% 
  # Top 3 worst profitability
  slice_head(n = 3)

quiet_underperformers
```
